---
title: "CHATON Trial Primary Analysis - CONSORT Participant Flow"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---

# Summary

Track participant flow in the CHATON trial.

Summarize patient numbers by treatment group and phase to include in a CONSORT flow diagram.

# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here("R/0-CHATON-primary-Config.R"))
options(knitr.kable.NA = '')
```

# Load the data

Load the final population tracking dataset. This file was created by `0-CHATON-primary-data-processing.R`.

```{r load the data}
#---------------------------------
# load the pre-processed data
#---------------------------------
d <- read_rds(here("data","CHATON_full_dataset_public.rds"))

# prescreen dataset
inelig <- read_rds(here("data","CHATON_inelig_dataset_public.rds"))
```

# CONSORT flow

Participant flow for participants randomized and followed through their 14 days and 6-month visit.

```{r}
#---------------------------------
# 0. Prescreening
#---------------------------------
prescreen <- inelig %>%
  summarise(N_screen = n() + nrow(d),
            n_excl =n(),
            n_young = sum(too_young),
            n_old = sum(too_old),
            n_unreldob = sum(dob_unreliable),
            n_nooral = sum(no_oral_feeding),
            n_nostay6months = sum(not_stay6months),
            n_allergy = sum(allergy),
            n_refuse = sum(refuse)
            ) %>%
  t() %>%
  as.data.frame() %>%
  rename(Overall = V1) %>%
  mutate(Overall = as.character(Overall))

#----------------
# 1. Allocated to treatment and administered
#----------------
c_0 <- d %>%
  mutate(reason = ifelse(notadministered==1, "Vomited", 
                         ifelse(notadministered==2, "Refusal", 
                                "Other"))) %>%
  group_by(tr) %>%
  summarise(n_randomized = n(),
            n_administered = sum(isadministered == 1, na.rm = T),
            n_refusal=sum(reason=="Refusal", na.rm = T),
            n_vomited=sum(reason=="Vomited", na.rm = T),
            n_misstxform=sum(is.na(reason) & is.na(isadministered))) %>% 
  adorn_totals("row")

#----------------
# 2. 14 day 
#----------------
c_14 <-d %>%
  group_by(tr) %>%
  summarise(n_selected_14d = sum(selected14dcheck == 1, na.rm = T),
            n_analyzed_14d = sum(vital_14d %in% c("Alive", "Death")),
            n_ltfu_14d = n_selected_14d - n_analyzed_14d
            ) %>% 
  adorn_totals("row")

#----------------
# 3. 6 Month 
#----------------
c_180 <- d %>% 
  group_by(tr) %>%
  summarise(n_analyzed_6m = sum(outcome %in% c("Alive", "Death") &
                                  inwindow == 1),
            n_ltfu_6m = sum(outcome == "Unknown" | is.na(outcome)), 
            n_outwindow_6m = sum(outcome %in% c("Alive", "Death") & 
                                   inwindow == 0, na.rm = T)
            ) %>% 
  adorn_totals("row")

consort <- bind_rows(
  prescreen,
  c_0 %>% 
  full_join(c_14, by = "tr" ) %>%
  full_join(c_180, by = "tr" ) %>%
  # transpose
  t() %>% 
  # convert 1st row (tx) to colnames 
  row_to_names(1) %>%
  # convert matrix output of transpose to df 
  as.data.frame()%>% 
  # rename Total appropriately
  rename(Overall = Total)
  ) %>%
  select(Azithromycin,Placebo, Overall) %>%
  # set column name to empty string
  rownames_to_column(" ") %>% 
  # rename variables for presentation
  mutate(` ` = c("Assessed for eligibility",
                 "Excluded",
                 "Too young",
                 "Too old",
                 "Birth date source unreliable",
                 "No oral feeding",
                 "Not staying 6 months",
                 "Allergy", 
                 "Refused",
                 "Randomized",
                 "Administered",
                 "Refusal",
                 "Vomited",
                 "Missing treatment form",
                 "Selected",
                 "Analyzed",
                 "Lost to follow up/unknown vital status",
                 "Analyzed",
                 "Lost to follow up/unknown vital status",
                 "Measured out of window")
         )

consort %>% 
  knitr::kable(format = "html", digits = 0, format.args = list(big.mark = ",")
               ) %>%
  kable_styling(bootstrap_options = "striped") %>%
  pack_rows("Prescreen", 1, 9) %>%
  pack_rows("Enrollment", 10, 14) %>%    
  pack_rows("14 days follow-up", 15, 17) %>%  
  pack_rows("6 month follow-up", 18, 20) 

```

_Note: The measurement window for 6 months is age 4 months to 9 months (age 120 days to 270 days). Children measured before age 120 days or after age 270 days were excluded from the analysis._  

# Session Info
```{r session info}
sessionInfo()
```

