---
title: "CHATON Trial Primary Analysis - Mortality Sensitivity (Supplemental Material)"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---
# Summary

The report summarizes sensitivity analysis restricting only to infants residing in communities not enrolled in the CHAT trial.

# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here("R/0-CHATON-primary-Config.R"))
options(knitr.kable.NA = '')
```

# Load the data

Load the final population tracking dataset. This file was created by `0-CHATON-primary-data-processing.R`.

```{r load the data}
#---------------------------------
# load the pre-processed data
#---------------------------------
d <- read_rds(here("data","CHATON_full_dataset_public.rds")) %>%
  mutate(tr = factor(tr, levels = c("Placebo","Azithromycin")))
```

# Vital status at 6-months, in-window, excluded all CHAT villages

```{r}
d %>%
  filter(outcome %in% c("Alive", "Death") & inwindow == 1) %>%
  filter(village_in_chat != "In CHAT") %>%
  mutate(outcome = factor(outcome, 
                          levels = c('Alive', 
                                     'Death'))
         ) %>%
  set_variable_labels(outcome = "Vital Status at 6 months") %>%
  table1(~ outcome | tr, data = .)
```

# Vital status at 6-months, in-window, excluded CHAT villages with Azithromycin

```{r}
d %>%
  filter(outcome %in% c("Alive", "Death") & inwindow == 1) %>%
  filter(chat_tr == "Placebo/Not treated") %>%
  mutate(outcome = factor(outcome, 
                          levels = c('Alive', 
                                     'Death'))
         ) %>%
  set_variable_labels(outcome = "Vital Status at 6 months") %>%
  table1(~ outcome | tr, data = .)
```

# 1. Excluded all CHAT villages

## Hazard ratio permutation distribution

```{r}
#-------------------------------
# use the final child status
# by 6 months 
#
# set children with Unknown status to missing
#
# They will thus be excluded
# from the HR model estimates,
# be will be _included_ in the
# permutation test
#-------------------------------
model_df <- d %>%
  filter(outcome %in% c("Alive", "Death") & inwindow == 1) %>%
  filter(village_in_chat != "In CHAT") %>%
  select(tr, outcome, death)
#-------------------------------
# run the model
#-------------------------------
mod_empir <- glm(death ~ tr, data = model_df, family = binomial(link = cloglog))

# save a model summary
mod_summary <- summary(mod_empir) 

# save the log HR and its SE
coef_empir <- mod_summary$coef[2,1]
se_empir <- mod_summary$coef[2,2]

# calculate the upper and lower limits on the HR ratio (coefficient from the model)
upper_empir <- exp(coef_empir+(1.96*se_empir)) %>% round(3)
lower_empir <- exp(coef_empir-(1.96*se_empir)) %>% round(3)
coef_empir_exp <- exp(coef_empir) %>% round(3)
```

```{r}
# Predicted log risk
cloglog_az <- mod_empir$coefficients[1] + mod_empir$coefficients[2]
cloglog_pl <- mod_empir$coefficients[1]

# Convert log risk to probabilities
p_az <- 1 - exp(-exp(cloglog_az))
p_pl <- 1 - exp(-exp(cloglog_pl))

# Compute the absolute difference
p_diff <- p_az - p_pl 

# Compute the standard error of the difference using the delta method
var_diff <- (p_az * (1 - p_az) / nrow(model_df)) + (p_pl * (1 - p_pl) / nrow(model_df)) 
se_diff <- sqrt(var_diff)

# Compute the 95% confidence interval for the absolute difference
upper_diff <- p_diff + 1.96 * se_diff
lower_diff <- p_diff - 1.96 * se_diff
```

```{r}
#-------------------------------
# permute the treatment allocation and rerun the model to 
# estimate distribution of the model coefficient under the Null
#-------------------------------
vitalstatus <- model_df$death
tx_alloc <- model_df$tr

N <- 10000
null_vec <- foreach(simi = 1:N, .combine = rbind) %dopar% {
  set.seed(simi)
  
  # pull random numbers the length of treatment vector from uniform distribution 0-1
  reshuffle_vec <- runif(length(tx_alloc))
  
  # reshuffle the treatment allocation based on random numbers generated
  shuffled_tx <- tx_alloc[order(reshuffle_vec)]
 
   # subset to non-missing obs
  vitalstatusi <- vitalstatus[!is.na(vitalstatus)]
  shuffled_tx <- shuffled_tx[!is.na(vitalstatus)]
  
  # run the model on shuffled allocation
  modi <- glm(vitalstatusi ~ shuffled_tx, family = binomial(link = "cloglog"))
  
  # extract and save the coefficient
  return(modi$coefficients[2])
  
}

#-------------------------------
# calculate the 2-sided 
# permutation test p-value
#-------------------------------
p_value <- sum(abs(null_vec) >= abs(coef_empir)) / length(null_vec)

#-------------------------------
# plot the null distribution 
# with the empirical value
#-------------------------------
df <- data.frame(null_vec=null_vec)
ggplot(df, aes(x=null_vec)) + 
  geom_density(color = cbpal[2], fill=cbpal[2], alpha=0.5) +
  geom_vline(xintercept = coef_empir, color="black") +
  xlab("Null distribution of log(HR)") +
  theme_minimal()
```

## Mortality comparison

```{r}
#-------------------------------
# summary of children and deaths
#-------------------------------
df <- model_df %>%
  group_by(tr) %>%
  # tabulate deaths and total participants by arm
  summarise(N_ch=n(), N_d=sum(outcome =="Death")) %>%
  # calculate cumulative mortality
  mutate(d_percent=percent(N_d/N_ch, accuracy = .01),
         ) %>%
  pivot_wider(names_from=tr, values_from=c("N_ch", "N_d", "d_percent")) %>%
  select(N_ch_Azithromycin, N_d_Azithromycin, d_percent_Azithromycin,
         N_ch_Placebo, N_d_Placebo, d_percent_Placebo
         )

# prepare the hazard ratio from the model and p-value
df_hr <- data.frame(hr=paste0(sprintf("%1.2f",coef_empir_exp), 
                              " (", sprintf("%1.2f",lower_empir), ", ", 
                              sprintf("%1.2f",upper_empir), ")"), 
                    rd = paste0(sprintf("%1.2f",p_diff*100), 
                              "% (", sprintf("%1.2f",lower_diff*100), "%, ", 
                              sprintf("%1.2f",upper_diff*100), "%)"),
                    pvalue=p_value)

# bind everything together
df <- bind_cols(data.frame(outcome="Death by 6 month visit"), df, df_hr)

# render table
knitr::kable(df, 
             digits = 2,
             col.names = c("Outcome", "N Children", "N Deaths", "%", "N Children",
                     "N Deaths", "%", "HR (95% CI)", "Rate Diff (95% CI)", "P-value*"),
             caption = "Cumulative incidence of mortality by 6 months and hazard ratio (HR) for comparison between arms estimated using binomial regression with a complementary log-log link."
      ) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3, "Placebo" = 3, " " = 3)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)"))

```

# 2. Excluded CHAT villages treated with Azithromycin

## Hazard ratio permutation distribution

```{r}
#-------------------------------
# use the final child status
# by 6 months 
#
# set children with Unknown status to missing
#
# They will thus be excluded
# from the HR model estimates,
# be will be _included_ in the
# permutation test
#-------------------------------
model_df <- d %>%
  filter(outcome %in% c("Alive", "Death") & inwindow == 1) %>%
  filter(chat_tr == "Placebo/Not treated") %>%
  select(tr, outcome, death)
#-------------------------------
# run the model
#-------------------------------
mod_empir <- glm(death ~ tr, data = model_df, family = binomial(link = cloglog))

# save a model summary
mod_summary <- summary(mod_empir) 

# save the log HR and its SE
coef_empir <- mod_summary$coef[2,1]
se_empir <- mod_summary$coef[2,2]

# calculate the upper and lower limits on the HR ratio (coefficient from the model)
upper_empir <- exp(coef_empir+(1.96*se_empir)) %>% round(3)
lower_empir <- exp(coef_empir-(1.96*se_empir)) %>% round(3)
coef_empir_exp <- exp(coef_empir) %>% round(3)
```

```{r}
# Predicted log risk
cloglog_az <- mod_empir$coefficients[1] + mod_empir$coefficients[2]
cloglog_pl <- mod_empir$coefficients[1]

# Convert log risk to probabilities
p_az <- 1 - exp(-exp(cloglog_az))
p_pl <- 1 - exp(-exp(cloglog_pl))

# Compute the absolute difference
p_diff <- p_az - p_pl 

# Compute the standard error of the difference using the delta method
var_diff <- (p_az * (1 - p_az) / nrow(model_df)) + (p_pl * (1 - p_pl) / nrow(model_df)) 
se_diff <- sqrt(var_diff)

# Compute the 95% confidence interval for the absolute difference
upper_diff <- p_diff + 1.96 * se_diff
lower_diff <- p_diff - 1.96 * se_diff
```

```{r}
#-------------------------------
# permute the treatment allocation and rerun the model to 
# estimate distribution of the model coefficient under the Null
#-------------------------------
vitalstatus <- model_df$death
tx_alloc <- model_df$tr

N <- 10000
null_vec <- foreach(simi = 1:N, .combine = rbind) %dopar% {
  set.seed(simi)
  
  # pull random numbers the length of treatment vector from uniform distribution 0-1
  reshuffle_vec <- runif(length(tx_alloc))
  
  # reshuffle the treatment allocation based on random numbers generated
  shuffled_tx <- tx_alloc[order(reshuffle_vec)]
 
   # subset to non-missing obs
  vitalstatusi <- vitalstatus[!is.na(vitalstatus)]
  shuffled_tx <- shuffled_tx[!is.na(vitalstatus)]
  
  # run the model on shuffled allocation
  modi <- glm(vitalstatusi ~ shuffled_tx, family = binomial(link = "cloglog"))
  
  # extract and save the coefficient
  return(modi$coefficients[2])
  
}

#-------------------------------
# calculate the 2-sided 
# permutation test p-value
#-------------------------------
p_value <- sum(abs(null_vec) >= abs(coef_empir)) / length(null_vec)

#-------------------------------
# plot the null distribution 
# with the empirical value
#-------------------------------
df <- data.frame(null_vec=null_vec)
ggplot(df, aes(x=null_vec)) + 
  geom_density(color = cbpal[2], fill=cbpal[2], alpha=0.5) +
  geom_vline(xintercept = coef_empir, color="black") +
  xlab("Null distribution of log(HR)") +
  theme_minimal()
```

## Mortality comparison

```{r}
#-------------------------------
# summary of children and deaths
#-------------------------------
df <- model_df %>%
  group_by(tr) %>%
  # tabulate deaths and total participants by arm
  summarise(N_ch=n(), N_d=sum(outcome =="Death")) %>%
  # calculate cumulative mortality
  mutate(d_percent=percent(N_d/N_ch, accuracy = .01),
         ) %>%
  pivot_wider(names_from=tr, values_from=c("N_ch", "N_d", "d_percent")) %>%
  select(N_ch_Azithromycin, N_d_Azithromycin, d_percent_Azithromycin,
         N_ch_Placebo, N_d_Placebo, d_percent_Placebo
         )

# prepare the hazard ratio from the model and p-value
df_hr <- data.frame(hr=paste0(sprintf("%1.2f",coef_empir_exp), 
                              " (", sprintf("%1.2f",lower_empir), ", ", 
                              sprintf("%1.2f",upper_empir), ")"), 
                    rd = paste0(sprintf("%1.2f",p_diff*100), 
                              "% (", sprintf("%1.2f",lower_diff*100), "%, ", 
                              sprintf("%1.2f",upper_diff*100), "%)"),
                    pvalue=p_value)

# bind everything together
df <- bind_cols(data.frame(outcome="Death by 6 month visit"), df, df_hr)

# render table
knitr::kable(df, 
             digits = 2,
             col.names = c("Outcome", "N Children", "N Deaths", "%", "N Children",
                     "N Deaths", "%", "HR (95% CI)", "Rate Diff (95% CI)", "P-value*"),
             caption = "Cumulative incidence of mortality by 6 months and hazard ratio (HR) for comparison between arms estimated using binomial regression with a complementary log-log link."
      ) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3, "Placebo" = 3, " " = 3)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)"))

```

# Session Info

```{r}
sessionInfo()
```
