---
title: "CHATON Trial Primary Analysis - Adverse Events"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Track participant flow in the CHATON trial.

Summarize patient numbers by treatment group and phase to include in a CONSORT flow diagram.

# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here("R/0-CHATON-primary-Config.R"))
options(knitr.kable.NA = '')
```

# Load the data

Load the final population tracking dataset. This file was created by `0-CHATON-primary-data-processing.R`.

```{r load the data}
#---------------------------------
# load the pre-processed data
#---------------------------------
d <- read_rds(here("data","CHATON_full_dataset_public.rds")) %>%
  mutate(tr = factor(tr, levels = c("Placebo","Azithromycin")))
```


# SAEs at day 14

A random sample of 10% of children in the trial have been measured 14d later to assess serious adverse events (SAEs). Among participants measured at day 14 of follow-up, summarize the proportion with different SAEs.

Tabulation of any occurrence of each symptom by 14 days among study participants.

```{r}
#-------------------------------
# make the SAE variables factor
# variables for a table
# Any serious adverse event 
# Mortality within 14 days of treatment 
# Hospitalization within 28 days of treatment   
#-------------------------------
dsae <- d %>%
  filter(!is.na(childvital_14d)) %>%
  # re-code hospitalization - change 1/2 to 0/1
  mutate(hospitalized_14d = case_when(hospitalized_14d == 1 ~ 1, 
                                      hospitalized_14d == 2 ~ 0,
                                      TRUE ~ NA_real_)) %>%
  mutate(any_SAE = case_when(death_14d == 1 | hospitalized_14d == 1 ~ 1, 
                             death_14d == 0 & hospitalized_14d == 0 ~ 0,
                           TRUE ~ NA_real_)
         ) %>%
  mutate(
    any_sae = factor(any_SAE, levels=c(0,1), labels = c("No","Yes")),
    death = factor(death_14d, levels=c(0,1), labels = c("No","Yes")),
    hospitalized = factor(hospitalized_14d, levels=c(1,2), labels = c("Yes","No"))
         )
label(dsae$any_sae) <- "Any serious adverse event"
label(dsae$death) <- "Mortality within 14 days of treatment"
label(dsae$hospitalized) <- "Hospitalization within 14 days of treatment"

table1(~ any_sae + death + hospitalized | tr, data = dsae)
```

```{r SAE differences between arms}
#--------------------------
# pivot the SAE symptoms to
# long format
#--------------------------
sae_col <- c("any_SAE", "death_14d", "hospitalized_14d")
d_sael <- dsae %>% 
  group_by(id_public,tr) %>%
  select(id_public, tr, all_of(sae_col))%>%
  # reshape longer to make it easier to iterate the analysis
  pivot_longer(cols=sae_col,names_to = "sae_name",values_to = "sae") %>%
  # ensure treatment assignment is a factor
  # order sae_name according to the order we'd like in the table
  mutate(sae_name = factor(sae_name, 
                          levels = c("any_SAE", 
                                     "death_14d", 
                                     "hospitalized_14d"), 
                          labels = c("Any serious adverse event", 
                                     "Mortality within 14 days of treatment",
                                     "Hospitalization within 14 days of treatment")
         ))

#--------------------------
# summarize risk and risk
# difference between arms
# estimate the 95% CI using
# a linear binomial model
#--------------------------
sae_rd_ests <- foreach(saei = levels(d_sael$sae_name),.combine = rbind) %do% {
  # subset to the particular SAE
  di <- d_sael %>% filter(sae_name == saei)
  # estimate Ns by arm
  dns <- di %>%
    group_by(tr) %>%
    summarize(N = n(), n = sum(sae, na.rm = TRUE), .groups = "keep")
  
  #non-parametric bootstrap (1000 iterations) due to rare outcome
  set.seed(202304)
  aeboot <- foreach(booti = 1:1000, .combine = rbind) %do% {
    dii <- di[sample(1:nrow(di),replace=TRUE),]
    ia_N <- length(dii$sae[dii$tr=="Azithromycin"])
    ia_n <- sum(dii$sae[dii$tr=="Azithromycin"],na.rm=TRUE)
    ip_N <- length(dii$sae[dii$tr=="Placebo"])
    ip_n <- sum(dii$sae[dii$tr=="Placebo"],na.rm=TRUE)
    rd <- (ia_n/ia_N) - (ip_n/ip_N)
    return(rd)
    }
  pct_a <- dns$n[2]/dns$N[2]
  pct_p <- dns$n[1]/dns$N[1]
  rd <- pct_a - pct_p
  rd_ci <- quantile(aeboot,probs=c(0.025,0.975))
  res <- data.frame(sae_name = saei, 
                    N_a = dns$N[2], n_a = dns$n[2], pct_a = pct_a*100,
                    N_p = dns$N[1], n_p = dns$n[1], pct_p = pct_p*100,
                    rd, rd_se = NA_real_, rd_min95 = rd_ci[1], rd_max95 = rd_ci[2]) 
  res
}


#--------------------------
# print a summary table
#--------------------------
sae_rd_tab <- sae_rd_ests %>%
  mutate(npct_a = paste0(n_a," (",sprintf("%1.2f",pct_a),"%)"),
         npct_p = paste0(n_p," (",sprintf("%1.2f",pct_p),"%)"),
         rd95ci = paste0(sprintf("%1.2f",rd*100),"% (", sprintf("%1.2f",rd_min95*100),"% to ",sprintf("%1.2f",rd_max95*100),"%)")
         )
rownames(sae_rd_tab)<-NULL

knitr::kable(sae_rd_tab %>% select(sae_name,npct_a,npct_p,rd95ci),
             align = "lrrr",
             col.names = c("Serious Adverse Event", 
                          paste0("Azithromycin", " \n N=", unique(sae_rd_tab$N_a), " n(%)"),
                           paste0("Placebo", " \n N=", unique(sae_rd_tab$N_p), " n(%)"),
                           "RD (95% CI) *"),
             row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = TRUE) %>%
  footnote(symbol = c("RD: risk difference (azithromycin - placebo), in percentage points. The 95% CIs were estimated with a non-parametric bootstrap (1000 iterations) due to the small number of events."))
```


# Non-serious AEs at day 14

```{r}
#----------------------------------
# for participants measured at 
# day 14, limit the data to 
# symptoms measured then.
#
# then summarize the number and %
# who had each symptom
#----------------------------------

dae <- d %>%
  filter(!is.na(vital_14d)) %>%
  mutate(any_AE = case_when(symptoms_abdominalpain_14d == 1 | 
                              symptoms_vomiting_14d == 1 | 
                              symptoms_diarrhea_14d == 1 | 
                              symptoms_constipation_14d == 1 | 
                              symptoms_hemorrhoids_14d == 1 | 
                              symptoms_rash_14d == 1 ~ 1, 
                            symptoms_abdominalpain_14d == 0 & 
                             symptoms_vomiting_14d == 0 & 
                              symptoms_diarrhea_14d == 0 & 
                              symptoms_constipation_14d == 0 & 
                              symptoms_hemorrhoids_14d == 0 & 
                              symptoms_rash_14d == 0 ~ 0,
                           TRUE ~ NA_real_)
         ) %>%
  mutate(
    any_ae = factor(any_AE, levels=c(0,1), labels = c("No","Yes")),
    abdominal = factor(symptoms_abdominalpain_14d, levels=c(0,1), labels = c("No","Yes")),
    vomiting = factor(symptoms_vomiting_14d, levels=c(0,1), labels = c("No","Yes")),
    diarrhea = factor(symptoms_diarrhea_14d, levels=c(0,1), labels = c("No","Yes")),
    constipation = factor(symptoms_constipation_14d, levels=c(0,1), labels = c("No","Yes")),
    hemorrhoids = factor(symptoms_hemorrhoids_14d, levels=c(0,1), labels = c("No","Yes")),
    rash = factor(symptoms_rash_14d, levels=c(0,1), labels = c("No","Yes"))
         )
label(dae$any_ae) <- "Any non-serious adverse event"
label(dae$abdominal) <- "Abdominal pain"
label(dae$vomiting) <- "Vomiting"
label(dae$diarrhea) <- "Diarrhea"
label(dae$constipation) <- "Constipation"
label(dae$hemorrhoids) <- "Hemorrhoids"
label(dae$rash) <- "Rash"

table1(~ any_ae + abdominal + vomiting + diarrhea + constipation + hemorrhoids + rash | tr, data = dae )
```


```{r}
#--------------------------
# pivot the AE symptoms to
# long format
#--------------------------
ae_col <- c("any_ae","abdominal","vomiting","diarrhea","constipation","hemorrhoids","rash")
d_ael <- dae %>% 
  group_by(id_public,tr) %>%
  select(id_public,tr,ae_col) %>%
  # reshape longer to make it easier to iterate the analysis
  pivot_longer(cols=ae_col,names_to = "ae_name",values_to = "ae") %>%
  # ensure treatment assignment is a factor
  # order ae_name according to the order we'd like in the table
  mutate(ae_name = factor(ae_name, 
                          levels = c("any_ae",
                                     "abdominal","vomiting",
                                     "diarrhea","constipation",
                                     "hemorrhoids","rash"),
                          labels = c("Any non-serious adverse event", 
                                     "Abdominal pain", "Vomiting", 
                                     "Diarrhea", "Constipation", 
                                     "Hemorrhoids", "Rash")))

#--------------------------
# summarize risk and risk
# difference between arms
# estimate the 95% CI using
# a linear binomial model
#--------------------------
ae_rd_ests <- foreach(aei = levels(d_ael$ae_name),.combine = rbind) %do% {
  # subset to the particular AE
  di <- d_ael %>% filter(ae_name == aei)
  # estimate Ns by arm
  dns <- di %>% 
    group_by(tr) %>%
    summarize(N = n(), n = sum(ae == "Yes", na.rm = TRUE), .groups = "keep")
  # fit a linear binomial model. coefficient on tx estimates the RD
  fiti <- glm(ae ~ tr, data = di, family = binomial(link = "identity"))
  sumi <- summary(fiti)
  # get the estimate and its SE
  esti <- fiti$coefficients[2]
  sei <- sqrt(diag(sumi$cov.unscaled)[2])
  # return results
  res <- data.frame(ae_name = aei, 
                    N_a = dns$N[2], n_a = dns$n[2], pct_a = 100*dns$n[2]/dns$N[2],
                    N_p = dns$N[1], n_p = dns$n[1], pct_p = 100*dns$n[1]/dns$N[1],
                    rd = esti, rd_se = sei, rd_min95 = esti -1.96*sei, rd_max95 = esti + 1.96*sei)
  res
}

#--------------------------
# print a summary table
#--------------------------
ae_rd_tab <- ae_rd_ests %>%
  mutate(npct_a = paste0(n_a," (",sprintf("%1.1f",pct_a),"%)"),
         npct_p = paste0(n_p," (",sprintf("%1.1f",pct_p),"%)"),
         rd95ci = paste0(sprintf("%1.1f",rd*100),"% (", sprintf("%1.1f",rd_min95*100),"% to ",sprintf("%1.1f",rd_max95*100),"%)")
         )
rownames(ae_rd_tab) <- NULL

knitr::kable(ae_rd_tab %>% select(ae_name,npct_a,npct_p,rd95ci),
             align = "lrrr",
             col.names = c("Adverse event", 
                           paste0("Azithromycin N=", unique(ae_rd_ests$N_a), " n(%)"),
                           paste0("Placebo N=", unique(ae_rd_ests$N_p), " n(%)"),
                           "RD (95% CI) *"),
             row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped") %>%
  footnote(symbol = c("RD: risk difference (azithromycin - placebo), in percentage points."))
```

# Session Info

```{r}
sessionInfo()
```

