---
title: "CHATON Trial Primary Analysis - Secondary outcomes"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---
# Summary

The report summarizes secondary endpoints at 6 months in the CHATON trial.

# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here("R/0-CHATON-primary-Config.R"))
options(knitr.kable.NA = '')
```

# Load the data

Load the final population tracking dataset. This file was created by `0-CHATON-primary-data-processing.R`.

```{r load the data}
#---------------------------------
# load the pre-processed data
#---------------------------------
d <- read_rds(here("data","CHATON_full_dataset_public.rds")) %>%
  mutate(tr = factor(tr, levels = c("Placebo","Azithromycin")))
```

# Hospitalization

```{r}
# summarize Hospitalization at 6 month followup
d_hos <- d %>%
  mutate(hospitalized_6m = case_when(hospitalized_6m == 1 ~ "Yes", 
                                     hospitalized_6m == 2 ~ "No", 
                                     T ~ "Unknown", 
                                     ),
         hos_6m = case_when(hospitalized_6m == "Yes" ~ 1,
                            hospitalized_6m == "No" ~ 0, 
                            T ~ NA_real_, 
                                     )) %>%
  filter(inwindow == 1 & 
           hospitalized_6m %in% c("Yes", "No"))

d_hos %>%
  mutate(hospitalized_6m = factor(hospitalized_6m, levels = c("Yes", "No"))) %>%
  set_variable_labels(hospitalized_6m = "Any hospitalization") %>%
  table1(~ hospitalized_6m | tr, data = .)
```


## Hazard ratio permutation distribution

```{r}
#-------------------------------
# use the final child status
# by 6 months 
#
# set children with Unknown status to missing
#
# They will thus be excluded
# from the HR model estimates,
# be will be _included_ in the
# permutation test
#-------------------------------
model_df_hos <- d_hos %>%
  select(id_public, tr, hos_6m, hospitalized_6m) %>%
  mutate(tr = factor(tr, levels = c("Placebo", "Azithromycin")))
#-------------------------------
# run the model
#-------------------------------
mod_empir_hos <- glm(hos_6m ~ tr, data = model_df_hos, family = binomial(link = cloglog))

# save a model summary
mod_summary_hos <- summary(mod_empir_hos) 

# save the log HR and its SE
coef_empir_hos <- mod_summary_hos$coef[2,1]
se_empir_hos <- mod_summary_hos$coef[2,2]

# calculate the upper and lower limits on the HR ratio (coefficient from the model)
upper_empir_hos <- exp(coef_empir_hos+(1.96*se_empir_hos)) %>% round(3)
lower_empir_hos <- exp(coef_empir_hos-(1.96*se_empir_hos)) %>% round(3)
coef_empir_exp_hos <- exp(coef_empir_hos) %>% round(3)
```

```{r}
#-------------------------------
# permute the treatment allocation and rerun the model to 
# estimate distribution of the model coefficient under the Null
#-------------------------------
hos <- model_df_hos$hos_6m
tx_alloc <- model_df_hos$tr

N <- 10000
null_vec <- foreach(simi = 1:N, .combine = rbind) %dopar% {
  set.seed(simi)
  
  # pull random numbers the length of treatment vector from uniform distribution 0-1
  reshuffle_vec <- runif(length(tx_alloc))
  
  # reshuffle the treatment allocation based on random numbers generated
  shuffled_tx <- tx_alloc[order(reshuffle_vec)]
 
   # subset to non-missing obs
  hosi <- hos[!is.na(hos)]
  shuffled_tx <- shuffled_tx[!is.na(hos)]
  
  # run the model on shuffled allocation
  modi <- glm(hosi ~ shuffled_tx, family = binomial(link = "cloglog"))
  
  # extract and save the coefficient
  return(modi$coefficients[2])
  
}

#-------------------------------
# calculate the 2-sided 
# permutation test p-value
#-------------------------------
p_value_hos <- sum(abs(null_vec) > abs(coef_empir_hos)) / length(null_vec)

#-------------------------------
# plot the null distribution 
# with the empirical value
#-------------------------------
df <- data.frame(null_vec=null_vec)
ggplot(df, aes(x=null_vec)) + 
  geom_density(color = cbpal[2], fill=cbpal[2], alpha=0.5) +
  geom_vline(xintercept = coef_empir_hos, color="black") +
  xlab("Null distribution of log(HR)") +
  theme_minimal()
```

```{r}
#-------------------------------
# summary of children and deaths
#-------------------------------
df_hos <- model_df_hos %>%
  group_by(tr) %>%
  # tabulate deaths and total participants by arm
  summarise(N_ch=n(), N_h=sum(hospitalized_6m == "Yes")) %>%
  # calculate cumulative mortality
  mutate(h_percent=percent(N_h/N_ch, accuracy = .01)) %>%
  pivot_wider(names_from=tr, values_from=c("N_ch", "N_h", "h_percent")) %>%
  select(N_ch_Azithromycin, N_h_Azithromycin, h_percent_Azithromycin,
         N_ch_Placebo, N_h_Placebo, h_percent_Placebo
         )

# prepare the hazard ratio from the model and p-value
df_hr <- data.frame(hr=paste0(sprintf("%1.2f",coef_empir_exp_hos), 
                              " (", sprintf("%1.2f",lower_empir_hos), ", ", 
                              sprintf("%1.2f",upper_empir_hos), ")"), 
                    pvalue=p_value_hos)

# bind everything together
df <- bind_cols(data.frame(outcome="Any hospitalization"), df_hos, df_hr)

# render table
knitr::kable(df, 
             digits = 2,
             col.names = c("Outcome", "N Children", "N Hospitalization", "%", "N Children",
                     "N Hospitalization", "%", "HR (95% CI)", "P-value*"),
             caption = "Cumulative incidence of hospitalization by 6 months and hazard ratio (HR) for comparison between arms estimated using binomial regression with a complementary log-log link."
      ) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3, "Placebo" = 3, " " = 2)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)"))

```

# CSPS Visit

```{r}
# summarize clinic visit at 6 month followup
d_csps <- d %>%
  mutate(visitedcsps_6m = case_when(visitedcsps_6m == 1 ~ "Yes", 
                               visitedcsps_6m == 2 ~ "No", 
                               T ~ "Unknown", 
                                     ),
         csps_6m = case_when(visitedcsps_6m == "Yes" ~ 1,
                            visitedcsps_6m == "No" ~ 0, 
                            T ~ NA_real_, 
                                     ))%>%
  mutate() %>%
  filter(inwindow == 1 & 
           visitedcsps_6m %in% c("Yes", "No"))

d_csps %>%
  mutate(tr = factor(tr, levels = c("Azithromycin","Placebo")),
         visitedcsps_6m = factor(visitedcsps_6m, levels = c("Yes", "No"))
         ) %>%
  set_variable_labels(visitedcsps_6m = "Any CSPS Visit") %>%
  table1(~ visitedcsps_6m | tr, data = .)
```

## Hazard ratio permutation distribution

```{r}
#-------------------------------
# use the final child status
# by 6 months 
#
# set children with Unknown status to missing
#
# They will thus be excluded
# from the HR model estimates,
# be will be _included_ in the
# permutation test
#-------------------------------
model_df_csps <- d_csps %>%
  select(id_public, tr, csps_6m, visitedcsps_6m) %>%
  mutate(tr = factor(tr, levels = c("Placebo", "Azithromycin")))
#-------------------------------
# run the model
#-------------------------------
mod_empir_csps <- glm(csps_6m ~ tr, data = model_df_csps, family = binomial(link = cloglog))

# save a model summary
mod_summary_csps <- summary(mod_empir_csps) 

# save the log HR and its SE
coef_empir_csps <- mod_summary_csps$coef[2,1]
se_empir_csps <- mod_summary_csps$coef[2,2]

# calculate the upper and lower limits on the HR ratio (coefficient from the model)
upper_empir_csps <- exp(coef_empir_csps+(1.96*se_empir_csps)) %>% round(3)
lower_empir_csps <- exp(coef_empir_csps-(1.96*se_empir_csps)) %>% round(3)
coef_empir_exp_csps <- exp(coef_empir_csps) %>% round(3)
```

```{r}
#-------------------------------
# permute the treatment allocation and rerun the model to 
# estimate distribution of the model coefficient under the Null
#-------------------------------
csps <- model_df_csps$csps_6m
tx_alloc <- model_df_csps$tr

N <- 10000
null_vec <- foreach(simi = 1:N, .combine = rbind) %dopar% {
  set.seed(simi)
  
  # pull random numbers the length of treatment vector from uniform distribution 0-1
  reshuffle_vec <- runif(length(tx_alloc))
  
  # reshuffle the treatment allocation based on random numbers generated
  shuffled_tx <- tx_alloc[order(reshuffle_vec)]
 
   # subset to non-missing obs
  cspsi <- csps[!is.na(csps)]
  shuffled_tx <- shuffled_tx[!is.na(csps)]
  
  # run the model on shuffled allocation
  modi <- glm(cspsi ~ shuffled_tx, family = binomial(link = "cloglog"))
  
  # extract and save the coefficient
  return(modi$coefficients[2])
  
}

#-------------------------------
# calculate the 2-sided 
# permutation test p-value
#-------------------------------
p_value_csps <- sum(abs(null_vec) > abs(coef_empir_csps)) / length(null_vec)

#-------------------------------
# plot the null distribution 
# with the empirical value
#-------------------------------
df <- data.frame(null_vec=null_vec)
ggplot(df, aes(x=null_vec)) + 
  geom_density(color = cbpal[2], fill=cbpal[2], alpha=0.5) +
  geom_vline(xintercept = coef_empir_csps, color="black") +
  xlab("Null distribution of log(HR)") +
  theme_minimal()
```

```{r}
#-------------------------------
# summary of children and deaths
#-------------------------------
df_csps <- model_df_csps %>%
  group_by(tr) %>%
  # tabulate deaths and total participants by arm
  summarise(N_ch=n(), N_h=sum(visitedcsps_6m == "Yes")) %>%
  # calculate cumulative mortality
  mutate(h_percent=percent(N_h/N_ch, accuracy = .01)) %>%
  pivot_wider(names_from=tr, values_from=c("N_ch", "N_h", "h_percent")) %>%
  select(N_ch_Azithromycin, N_h_Azithromycin, h_percent_Azithromycin,
         N_ch_Placebo, N_h_Placebo, h_percent_Placebo
         )

# prepare the hazard ratio from the model and p-value
df_hr_csps <- data.frame(hr=paste0(sprintf("%1.2f",coef_empir_exp_csps), 
                              " (", sprintf("%1.2f",lower_empir_csps), ", ", 
                              sprintf("%1.2f",upper_empir_csps), ")"), 
                    pvalue=p_value_csps)

# bind everything together
df <- bind_cols(data.frame(outcome="Any CSPS visit"), df_csps, df_hr_csps)

# render table
knitr::kable(df, 
             digits = 2,
             col.names = c("Outcome", "N Children", "N CSPS visit", "%", "N Children",
                     "N CSPS visit", "%", "HR (95% CI)", "P-value*"),
             caption = "Cumulative incidence of CSPS visit by 6 months and hazard ratio (HR) for comparison between arms estimated using binomial regression with a complementary log-log link."
      ) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3, "Placebo" = 3, " " = 2)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)"))

``` 


```{r}
d_csps_reason <- d %>%
  mutate(malaria_6m = case_when(reasoncsps_malaria_6m == 1 ~ "Yes", 
                            reasoncsps_malaria_6m == 0 | visitedcsps_6m == 2  ~ "No", 
                            T ~ "Unknown" 
                                     ),
         pneumonia_6m = case_when(reasoncsps_pneumonia_6m == 1 ~ "Yes", 
                               reasoncsps_pneumonia_6m == 0 | visitedcsps_6m == 2 ~ "No", 
                               T ~ "Unknown" 
                                     ),
         diarrhea_6m = case_when(reasoncsps_diarrhea_6m == 1 ~ "Yes",
                            reasoncsps_diarrhea_6m == 1 | visitedcsps_6m == 2 ~ "No", 
                            T ~ "Unknown"
                                     ),
         reasoncsps_malaria_6m = case_when(reasoncsps_malaria_6m == 1 ~ 1,
                            reasoncsps_malaria_6m == 0 | visitedcsps_6m == 2 ~ 0, 
                            T ~ 99
                                     ),
         reasoncsps_pneumonia_6m = case_when(reasoncsps_pneumonia_6m == 1 ~ 1,
                            reasoncsps_pneumonia_6m == 0 | visitedcsps_6m == 2 ~ 0, 
                            T ~ 99
                                     ),
         reasoncsps_diarrhea_6m = case_when(reasoncsps_diarrhea_6m == 1 ~ 1,
                            reasoncsps_diarrhea_6m == 0 | visitedcsps_6m == 2 ~ 0, 
                            T ~ 99
                                     )
         ) %>%
  filter(inwindow == 1 & 
           visitedcsps_6m %in% c(1, 2)) 

d_csps_reason %>%
  mutate(tr = factor(tr, levels = c("Azithromycin","Placebo")),
         malaria_6m = factor(malaria_6m, levels = c("Yes", "No")),
         pneumonia_6m = factor(pneumonia_6m, levels = c("Yes", "No")),
         diarrhea_6m = factor(diarrhea_6m, levels = c("Yes", "No"))
         ) %>%
  set_variable_labels(malaria_6m = "CSPS Visit - malaria",
                      pneumonia_6m = "CSPS Visit - pneumonia",
                      diarrhea_6m = "CSPS Visit - diarrhea"
                      ) %>%
  table1(~ malaria_6m + pneumonia_6m + diarrhea_6m | tr, data = .)
```


## Hazard ratio permutation distribution

```{r}
result_df <- data.frame(
  outcome = NULL, 
  N_ch_Placebo = NULL, N_h_Placebo = NULL, h_percent_Placebo = NULL,
  N_ch_Azithromycin = NULL, N_h_Azithromycin = NULL, h_percent_Azithromycin = NULL,
  hr = NULL, 
  pvalue = NULL
)

model_df_csps <- d_csps_reason %>%
  select(id_public, tr, visitedcsps_6m, reasoncsps_malaria_6m, reasoncsps_pneumonia_6m, reasoncsps_diarrhea_6m)%>%
  mutate(tr = factor(tr, levels = c("Placebo", "Azithromycin")))
outcomes <- c("reasoncsps_malaria_6m", "reasoncsps_pneumonia_6m", "reasoncsps_diarrhea_6m")

for (o in outcomes) {
  temp <- model_df_csps %>% rename(outcome = o)
  #-------------------------------
  # run the model
  #-------------------------------
  mod_empir <- glm(outcome ~ tr, data = temp, family = binomial(link = "cloglog"))
  # save a model summary
  mod_summary <- summary(mod_empir) 

  # save the log HR and its SE
  coef_empir <- mod_summary$coef[2,1]
  se_empir <- mod_summary$coef[2,2]
  
  # calculate the upper and lower limits on the HR ratio (coefficient from the model)
  upper_empir <- exp(coef_empir+(1.96*se_empir)) %>% round(3)
  lower_empir <- exp(coef_empir-(1.96*se_empir)) %>% round(3)
  coef_empir_exp <- exp(coef_empir) %>% round(3)
  
  
  #-------------------------------
  # permute the treatment allocation and rerun the model to 
  # estimate distribution of the model coefficient under the Null
  #-------------------------------
  outcome <- temp$outcome
  tx_alloc <- temp$tr
  
  N <- 10000
  null_vec <- foreach(simi = 1:N, .combine = rbind) %dopar% {
    set.seed(simi)
    
    # pull random numbers the length of treatment vector from uniform distribution 0-1
    reshuffle_vec <- runif(length(tx_alloc))
    
    # reshuffle the treatment allocation based on random numbers generated
    shuffled_tx <- tx_alloc[order(reshuffle_vec)]
   
     # subset to non-missing obs
    outcomei <- outcome[!is.na(outcome)]
    shuffled_tx <- shuffled_tx[!is.na(outcome)]
    
    # run the model on shuffled allocation
    modi <- glm(outcomei ~ shuffled_tx, family = binomial(link = "cloglog"))
    
    # extract and save the coefficient
    return(modi$coefficients[2])
    
  }
  #-------------------------------
  # calculate the 2-sided 
  # permutation test p-value
  #-------------------------------
  p_value <- sum(abs(null_vec) > abs(coef_empir)) / length(null_vec)
  
  #-------------------------------
  # summary of children and deaths
  #-------------------------------
  summary_o <- temp %>%
    group_by(tr) %>%
    # tabulate deaths and total participants by arm
    summarise(N_ch=n(), N_h=sum(outcome == 1)) %>%
    # calculate cumulative mortality
    mutate(h_percent=percent(N_h/N_ch, accuracy = .01)) %>%
    pivot_wider(names_from=tr, values_from=c("N_ch", "N_h", "h_percent")) %>%
    select(N_ch_Placebo, N_h_Placebo, h_percent_Placebo,
           N_ch_Azithromycin, N_h_Azithromycin, h_percent_Azithromycin
           )
  
  # bind everything together
  df_o <- bind_cols(data.frame(outcome=o),
                  summary_o,
                  data.frame(hr=paste0(sprintf("%1.2f",coef_empir_exp), 
                                " (", sprintf("%1.2f",lower_empir), ", ", 
                                sprintf("%1.2f",upper_empir), ")"), 
                             pvalue=p_value))
  result_df <- bind_rows(result_df, df_o)
}
```

```{r}
#-------------------------------
# summary of children and deaths
#-------------------------------
result_df$outcome <- c("CSPS malaria", "CSPS pneumonia", "CSPS diarrhea")

result_df <- result_df %>%
  select(outcome,
         N_ch_Azithromycin, N_h_Azithromycin, h_percent_Azithromycin,
         N_ch_Placebo, N_h_Placebo, h_percent_Placebo,
         hr, pvalue
         )
# render table
knitr::kable(result_df, 
             digits = 2,
             col.names = c("Outcome", "N Children", "N CSPS visit", "%", "N Children",
                     "N CSPS visit", "%", "HR (95% CI)", "P-value*"),
             caption = "Cumulative incidence of CSPS visit by 6 months and hazard ratio (HR) for comparison between arms estimated using binomial regression with a complementary log-log link."
      ) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3, "Placebo" = 3, " " = 2)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)"))

```

# Session Info

```{r}
sessionInfo()
```
