---
title: "CHATON Trial Primary Analysis - Trial Population  (Supplemental Material)"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Joyce Hu, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "Data Cutoff: 2023-03-08. Report Updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../output"
                          )})
---
# Summary

The report summarizes age and sex distribution of the general population, which was derived from a biannual census that was conducted in the study area at the time of trial enrollment.

**Disclaimer:** This script runs off of data from another trial, CHAT, and will not run off of CHATON public replication data.

# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here("R/0-CHATON-primary-Config.R"))
options(knitr.kable.NA = '')
```

# Load the data

Load the final population tracking dataset. This file was created by `0-CHATON-primary-data-processing.R`.


```{r load the data}
#---------------------------------
# load the pre-processed data
#---------------------------------
d <- read_rds(here("data","CHATON_full_dataset_public.rds")) %>%
  mutate(tr = factor(tr, levels = c("Placebo","Azithromycin")))

#---------------------------------
# CHAT Village census population data
#
# **NOTE** This file is not currently
# available publicly, so this minor 
# script will not run in public 
# replication efforts
#---------------------------------
CHAT_child <- readRDS("~/Library/CloudStorage/Box-Box/Burkina Faso/CHAT/data/final/mortality/CHAT_child_census_dataset.rds")

```

# CHAT Population 

```{r}

my.render.cont <- function(x) {
  with(stats.apply.rounding(stats.default(x, ), digits = 2),
       c("",
         "median [Q1-Q3]" =
           sprintf(paste0("%s [",Q1,", %s]"), MEDIAN,Q3)))
}

# CHAT children who were eligible for CHATON
CHAT_child %>%
  filter(ageInMonths %in% 1:4) %>%
  filter(anneeNchild != 9999 | 
           moisNchild != 99 |
           jourNchild != 99) %>%
  select(analysis_ID, ageInMonths, startTime, 
         anneeNchild, moisNchild, jourNchild, childSex) %>%
  mutate(dob = as.Date(paste0(anneeNchild, "-", moisNchild, "-", jourNchild)),
         ageInDays = as.numeric(startTime - dob),
         ageInWeeks = round(ageInDays/7, 1),
         childsex = ifelse(childSex == 1, "Male", "Female")) %>%
  filter(ageInDays %in% 28:84 & !is.na(childsex)) %>%
  table1(~ ageInWeeks + childsex, data = ., 
         render.continuous=my.render.cont
  )
```

# CHATON Population 

```{r}
d %>%
  table1(~ ageinweek_bl + childsex, data = ., 
         render.continuous=my.render.cont
  )
```
# Session Info
```{r session info}
sessionInfo()
```
